Forecasting biodiversity in breeding birds using best practices

### 2017-12-08 manuscript/reviewer_reply.Rmd

*The reviewers and I agree that this manuscript is interesting, timely and well-written. I will add a few of my own observations here. First, and least important - normally submitted manuscripts do not hyphenate and do not right justify.*

>Thank you. We have removed the right justification and hyphenation.

*Next, in your figure 1, which forms the basis for comparison of model accuracy over time, I really expected you to choose data WITH a temporal pattern. After all, it looks to me like you have a time series in which species richness simply varies around a more or less stable average value, and the variation can be due to nothing (random), observer bias, and any number of species population trends (and combinations thereof). Thus, you have a situation of years of no trend, and use those data to predict no trend into the future? I would have thought you would choose data with a definite time trend that was already potentially correlated with, say, climate change, and then use hindcasting and forecasting to test whether the models accurately predict a trend into the "future"! To me, it is not surprising that models do not "...successfully turn spatial data into useful temporal predictions about biodiversity at decadal time-scales." But, I wonder if you chose data WITH a trend, you would have gotten other results.*

>>We added a "future directions" paragraph to address this. Non-directional changes, such as fluctuations from year-to-year, are forecastable in principle, although we agree that they may be harder. Our goal was to evaluate how well this worked for as many time series as possible. We agree that it would be interesting to examine the subset of locations that showed strong directional change and see if those are easier to forecast. We plan on coming back to investigate these questions at a later date.

@@

*Line 180 – Maybe also mention this is in the forecast package – as you do with gbm below?*

>We have added this.

*Line 183: Because these defaults may change over time, it’d be good to specify things like the ranges of AR / MA order that you searched over, which criteria was used to compare models, etc. And whether or not you included a drift parameter?*

>We agree that software can evolve over time and make methods non-repricable. We have added details of the auto.arima model in the supplement.

@@

>See comment above on hindcasting.

#### Methods:

*the authors perform their analyses on data that are collected from an incredibly dynamic time period. I wouldn’t expect species to have settled into a richness pattern given the incredibly disrupted nature of the last 32 years. Some strategies to try to handle these issues might be to train and predict at different temporal scales and to both forecast and hindcast the predictions. (e.g. train on 2004-2013 and try to predict 1982-1992 data + 1993-2003 data). If the authors were to make splits based on expected periods of dynamism, they could better tell whether the drivers that are included in the models are accurate predictors.*

>We agree that a comparison of model performance with different train/test splits would be interesting, and plan to make such comparisons in future work.

### 2017-12-08 manuscript/reviewer_reply.Rmd

*The reviewers and I agree that this manuscript is interesting, timely and well-written. I will add a few of my own observations here. First, and least important - normally submitted manuscripts do not hyphenate and do not right justify.*

>Thank you. We have removed the right justification and hyphenation.

*Next, in your figure 1, which forms the basis for comparison of model accuracy over time, I really expected you to choose data WITH a temporal pattern. After all, it looks to me like you have a time series in which species richness simply varies around a more or less stable average value, and the variation can be due to nothing (random), observer bias, and any number of species population trends (and combinations thereof). Thus, you have a situation of years of no trend, and use those data to predict no trend into the future? I would have thought you would choose data with a definite time trend that was already potentially correlated with, say, climate change, and then use hindcasting and forecasting to test whether the models accurately predict a trend into the "future"! To me, it is not surprising that models do not "...successfully turn spatial data into useful temporal predictions about biodiversity at decadal time-scales." But, I wonder if you chose data WITH a trend, you would have gotten other results.*

>We now address this in lines [[xx]]. We agree that non-directional changes, such as fluctuations from year-to-year, will generally be harder to predict, although it should still be possible to predict some of them. Some of the sites did show directional changes, and we agree that it would be interesting to examine this subset to see if those are easier to forecast. We plan on coming back to investigate these questions at a later date.

### 2017-12-08 manuscript/reviewer_reply.Rmd

*Line 183: Because these defaults may change over time, it’d be good to specify things like the ranges of AR / MA order that you searched over, which criteria was used to compare models, etc. And whether or not you included a drift parameter?*

>We agree that software can evolve over time and make methods non-repricable. We have added details of the auto.arima model in the supplement.

*Table 1 caption: without getting to more methods (below), I don’t understand the sentence ‘Environmental models were trained using all sites together, without information regarding which transects occurred at which site or during which year’.*

>We have clarified the Table 1 caption to be more clear about model covariates.

### 2017-12-08 manuscript/draft.Rmd

table_caption = "Six forecasting models. Single-site models were trained
site-by-site, without environmental data. Environmental models were trained
using all sites together, without information regarding which transects occurred
at which site or during which year. Most of the models were trained to predict
richness directly. This mirrors the standard application of these
techniques. Separate random forest SDMs were fit for each species and used to
predict the probability of that species occurring at each site. The
species-level probabilities at a site were summed to predict richness. The
mistnet JSDM was trained to predict the full species composition at each site,
and the number of species in its predictions was used as an estimate of
richness."

>>>>

table_caption = "Six forecasting models. Single-site models were trained
site-by-site, without environmental data. Environmental models were trained at
the continental scale, using only environmental variables (as opposed to site
or time series information) as predictors. Most of the models were trained to 
predict richness directly. This mirrors the standard application of these
techniques. Separate random forest SDMs were fit for each species and used to
predict the probability of that species occurring at each site. The
species-level probabilities at a site were summed to predict richness. The
mistnet JSDM was trained to predict the full species composition at each site,
and the number of species in its predictions was used as an estimate of
richness."
